\documentclass[aspectratio=169,english]{beamer}

\usepackage[english]{babel}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{pgfplots}
\usepackage{xifthen}
\usepackage{newunicodechar}
\usepackage{textcomp}
\usepackage[absolute,overlay]{textpos}
\usepackage{xparse}

\newunicodechar{°}{\textdegree}
\newunicodechar{µ}{\textmu}

\newcommand*{\ft}{\frametitle{\secname}}
\newcommand*{\ftsub}[1]{\frametitle{\secname\ - #1}}
\newcommand*{\itemp}{\item}
%\newcommand*{\itemp}{\pause\item}
\newcommand*{\+}{\texttt{+}}

\newcommand{\sectionframeimpl}[2]{
	\begin{frame}
	#1
	\begin{itemize}
	#2
	\end{itemize}
	\end{frame}
}
\NewDocumentCommand{\sectionframe}{s m m}{
	\IfBooleanTF{#1} {
		\section*{#2}
	} {
		\section{#2}
	}
	\sectionframeimpl{\ft}{#3}
}

\beamertemplatenavigationsymbolsempty

\mode<presentation>{
	\usetheme{Pittsburgh}
	\usecolortheme{seagull}
}

\begin{document}

\title{Rusticl}
\subtitle{OpenCL implemented in Rust}
\author{Karol Herbst}

\begin{frame}
\titlepage
\end{frame}

\sectionframe{State of OpenCL in Mesa - Clover}{
	\itemp Accepted in 2012 into Mesa
	\itemp Written in C++
	\itemp Targets LLVM (AMD) or NIR
	\itemp some development still happening (around 60 patches/year)
	\itemp Behaves differently compared to st/mesa
	\begin{itemize}
		\itemp $PIPE\_SHADER\_IR\_NIR\_SERIALIZED$
		\itemp Dynamic pipeloader (only user)
		\itemp $pipe\_compute\_state$::$req\_local\_mem$
		\itemp $pipe\_grid\_info$::$input$ for kernel inputs
	\end{itemize}
	\itemp Few users, poor support in projects using CL
}

\sectionframe{Motivation}{
	\itemp Rust sounded like a lot of fun
	\itemp Is Rust inside Mesa feasible and sustainable?
	\itemp Is implementing existing C APIs possible in Rust?
	\itemp What are the challenges working with C bindings?
	\itemp There are actually a lot of people asking about OpenCL.
}

\sectionframe{Why use Rust?}{
	\itemp If the compiler complains you are doing it wrong, period
	\itemp Very pleasant developing experience
	\itemp Enforces Thread and Memory Safety by design
	\itemp The threading API is a joy
}

\sectionframe{Rusticl}{
	\itemp Merged into Mesa in Sept. 2022
	\itemp A modern OpenCL 3.0 gallium frontend written in Rust
	\itemp Entirely SPIR-V and NIR based
	\itemp Tries to be as close as possible to st/mesa
	\itemp Low API overhead
	\itemp Images are supported!
	\itemp OpenCL conformant (Intel 12th gen and radeonsi)
	\itemp Worker thread architecture, thread private pipe\_context
}

\sectionframe{Meson}{
	\itemp Nice integration into existing build system
	\itemp Rust support good enough
	\itemp No Third-party Crate support yet
}

\sectionframe{Supported Drivers}{
	\itemp Iris
	\itemp llvmpipe
	\itemp panfrost (WIP?)
	\itemp radeonsi (WIP)
	\itemp nouveau (WIP)
	\itemp What's the minimum level of support to enable it by default?
}

\sectionframe{Supported Software}{
	\itemp Darktable
	\itemp Gimp
	\itemp Anything not using OpenCL extensions.. probably
}

\sectionframe{What is required from drivers}{
	\itemp Add it inside $src/gallium/targets/rusticl/meson.build$
	\itemp OpenGL Compute ($PIPE\_CAP\_COMPUTE$)
	\itemp $PIPE\_SHADER\_IR\_NIR$
	\itemp some $PIPE\_COMPUTE\_CAPs$
	\begin{itemize}
		\itemp $ADDRESS\_BITS$
		\itemp $MAX\_GLOBAL\_SIZE$
		\itemp $MAX\_MEM\_ALLOC\_SIZE$
		\itemp $SUBGROUP\_SIZE$
	\end{itemize}
	\itemp $pipe\_context$::$set\_global\_binding$
	\itemp for FULL\_PROFILE
	\begin{itemize}
		\itemp 128 textures
		\itemp 64 images
		\itemp 16 samplers
	\end{itemize}
	\itemp Function calling in backend compilers
}

\sectionframe{What is required from drivers Pt. 2}{
	\itemp $load/store\_global$
	\itemp For $resource\_map$ (\href{https://gitlab.freedesktop.org/mesa/mesa/-/merge_requests/18793}{!18793})
	\begin{itemize}
		\itemp $UNSYNCHRONIZED\ |\ COHERENT\ |\ PERSISTENT$
		\itemp on UMA: $UNSYNCHRONIZED\ |\ DIRECTLY$
	\end{itemize}
	\itemp \href{https://gitlab.freedesktop.org/mesa/mesa/-/merge_requests/18581}{!18581}
	\begin{itemize}
		\itemp $PIPE\_CAP\_SHAREABLE\_SHADER$
		\itemp $nir\_intrinsic\_load\_kernel\_input$ loading from cb0 or $lower\_uniforms\_to\_ubo$
		\itemp $pipe\_grid\_info.variable\_shared\_mem$
	\end{itemize}
}

\sectionframe{Issues with Bindings}{
	\itemp $static\ inline$
	\itemp unsigned instead of enum types
	\itemp bitfield enums
	\itemp forward declared OpenCL types
}

\sectionframe{Issues with other stuff}{
	\itemp gallium very much 32 bit specific
	\begin{itemize}
		\itemp buffers can only have a 31 bit size, 2GB
		\itemp ...
	\end{itemize}
	\itemp $pipe\_transfer$ can't be wrapped nicely
	\itemp OpenCLs memory API
	\itemp IDE integration
}

\sectionframe{TODOs}{
	\itemp Reduction of Kernel launch overhead !18581
	\itemp Proper implementation of $clGetKernelWorkGroupInfo$
	\itemp Improve utilization of the GPU
	\itemp Emitting less fences
	\itemp More API Validation
	\itemp More optional OpenCL features and extensions
	\itemp Function calling
	\itemp Make OpenCL "just work"
	\itemp maybe SyCL?
}

\sectionframe{Q\&A}{
	\itemp Any questions?
}

\end{document}